% SPDX-FileCopyrightText: 2026 Alan Szmyt
% SPDX-License-Identifier: Apache-2.0

\section{Experience}

\resumeentry{Software Engineer, Associate Staff}{MIT Lincoln Laboratory}{Lexington, MA}{Apr 2019 -- Aug 2025}

\begin{itemize}
\item Led development of a mobile tracking system for USAID to monitor commodities in low-connectivity environments, successfully field-testing pilot shipments to Africa and enabling end-to-end supply chain visibility.
\item Delivered a humanitarian deconfliction notification platform enabling analysts to visualize no-strike zones and submit secure, auditable reports, contributing to a follow-on effort that secured approximately \$1M in additional funding.
\item Built a counter-human trafficking investigation tool that enabled law enforcement analysts to triage large volumes of digital evidence and reduce time to actionable leads.
\item Developed externally accessible geospatial decision-support applications integrating shared data hubs, interactive mapping, and domain-specific datasets to improve situational awareness and analytical decision-making.
\item Contributed to NATO disaster-response exercises by extending field-ready mobile capabilities, including hazard visualization, area-of-interest management, and real-time location reporting across multinational operations.
\item Designed reproducible development environments and optimized Android (Gradle, Java/Kotlin) and modern web (Vite/ESBuild) build pipelines, improving onboarding speed, build reproducibility, and developer feedback cycles.
\item Developed full-stack geospatial systems for government and research sponsors, integrating data engineering, visualization, and automation for mission-critical analysis and field operations.
\end{itemize}

\vspace{0.5em}

\resumeentry{Founder \& Systems Architect}{Incompris LLC}{Wilmington, MA}{2023 -- Present}

\begin{itemize}
\item Founded and operate an independent creative-technology studio focused on AI-assisted workflow systems, deterministic publishing pipelines, and multimodal content automation.
\item Architected a specification-driven publishing platform that transforms canonical creative assets into reproducible multi-format artifacts (PDF, CBZ, WebP, AVIF, and LaTeX) through staged workflows and validation pipelines.
\item Designed AI-native workflow architectures that treat language models as bounded system components with explicit contracts, synchronization checkpoints, and reproducible execution paths.
\item Built multimodal content-generation pipelines integrating LLMs, image generation, video automation, metadata processing, and publication tooling into cohesive end-to-end workflows.
\item Developed local-first development environments and CI-backed automation systems using Docker, GitHub Actions, and reproducible infrastructure patterns to accelerate development and artifact validation.
\item Authored reflector, an independent research publication exploring synchronization architectures, recursive AI-assisted software engineering workflows, and human-AI governance systems.
\end{itemize}
